---
title: "for benchmark"
author: "Richel Bilderbeek"
date: "`r Sys.Date()`"
output: rmarkdown::pdf_document
vignette: >
  %\VignetteIndexEntry{for benchmark}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette aims to measure the alleged slowness
of for loops.

Here we have a non-trivial function:

```{r}
complex_function <- function(x) {
  out <- sqrt(x)
  out
}
```

Here we create two functions that work on a big vector,
with and without a for loop:

```{r}
f_for <- function() {
  values <- 1:100000
  n_values <- length(values)
  for (i in seq(1, n_values)) {
    values[i] <- complex_function(values[i])
  }
  values
}

f_no_for <- function() {
  values <- 1:100000
  values <- vapply(values, FUN = complex_function, FUN.VALUE = 0.0)
  values
}
```

They must give exactly the same results.

```{r}
testit::assert(f_for() == f_no_for())
```

Here we run a benchmark, comparing their speeds:

```{r}
tm <- microbenchmark::microbenchmark(f_for, f_no_for, times = 100000)
```

Plot as boxplot:

```{r fig.width = 7, fig.height = 7}
graphics::boxplot(tm)
```

View the raw results as table:

```{r}
knitr::kable(summary(tm))
```

Conclusion: when not growing vectors, using a for-loop
has little effect on run-time speed, if a non-trivial
function is used. Already the `sqrt` function is
non-trivial enough to show little effect.
