---
title: "is_distributed_normally"
author: "Richel Bilderbeek"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{is_distributed_normally}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette demonstrates the `is_distributed_normally` function.

```{r}
library(ribir)
```

Here we create a normally distributed collection of values:

```{r fig.width = 7, fig.height = 7}
nd <- rnorm(n = 1000, mean = 0.0, sd = 1.0)
plot(nd)
```

As we know that they are normally distributed, 
`is_distributed_normally` should return true:

```{r}
testit::assert(is_distributed_normally(nd))
```

Now we create a non-normal collection of values:

```{r fig.width = 7, fig.height = 7}
nnd <- runif(n = 1000, min = 0.0, max = 1.0)
plot(nnd)
```

As we know that they are *not* normally distributed, 
`is_distributed_normally` should return false:

```{r}
testit::assert(!is_distributed_normally(nnd))
```

## Use of `dplyr`

Here we'll try to do something smart:

```{r}
df_nd  <- data.frame(
  method = rep("nd" , times = length(nd)), x = nd
)

df_nnd <- data.frame(
  method = rep("nnd", times = length(nnd)), x = nnd
)
df <- rbind(df_nd, df_nnd)
df$method <- as.factor(df$method)

library(dplyr)
df_new <- df %>% 
  group_by(method) %>% 
  summarise(is_distributed_normally(x))

knitr::kable(df_new)
```
