---
title: "How to get the average nLTT plot"
author: "Richel Bilderbeek"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to get the average nLTT plot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Calculating the average nLTT plot of multiple phylogenies
is not a trivial tasks. The function `get_average_nltt`
does exactly this.


## Examples

For all examples, the following R code must be run:

```{r}
library(ape)
library(nLTT)
library(testit)
```

### Example: Easy trees

Create two easy trees:

```{r}
newick1 <- "((A:1,B:1):2,C:3);"
newick2 <- "((A:2,B:2):1,C:3);"
phylogeny1 <- ape::read.tree(text = newick1)
phylogeny2 <- ape::read.tree(text = newick2)
```

There are very similar. `phylogeny1` has short tips:

```{r}
plot(phylogeny1)
add.scale.bar()
```

This can be observed in the nLTT plot:

```{r}
nLTT.plot(phylogeny1, ylim = c(0,1))
```

As a collection of timepoints:

```{r}
ribir::get_phylogeny_nltt_matrix(phylogeny2)
```

`phylogeny2` has longer tips:

```{r}
plot(phylogeny2)
add.scale.bar()
```

Also this can be observed in the nLTT plot:

```{r}
nLTT.plot(phylogeny2, ylim = c(0,1))
```

As a collection of timepoints:

```{r}
ribir::get_phylogeny_nltt_matrix(phylogeny2)
```


The average nLTT plot should be somewhere in the middle.

It is constructed from strecthed nLTT matrices:

```{r}
ribir::stretch_nltt_matrix(ribir::get_phylogeny_nltt_matrix(phylogeny1), dt = 0.20, step_type = "upper")
```

```{r}
ribir::stretch_nltt_matrix(ribir::get_phylogeny_nltt_matrix(phylogeny2), dt = 0.20, step_type = "upper")
```



```{r}
ribir::get_average_nltt_matrix(c(phylogeny1,phylogeny2), dt = 0.20)
```

Or as a plot

```{r}
ribir::get_average_nltt(c(phylogeny1,phylogeny2), dt = 0.20, plot_nltts = TRUE)
```
